---
title: "Notes10-31-2024"
output: html_document
date: "2024-10-31"
---
install.packages("usethis")
library(usethis)
install.packages("devtools")
library(devtools)

ALL THIS CODE SHOULD ONLY BE LOADED IN THE CONSOLE TAB
```{r}
#create_package(".") makes the package in the location youa re currently at
this should give you a Namespace title README and Vignettes
##install() installs our project



```
Click the description file under project
it now looks liek this
Package: stack.bars
Title: takes data and creates a staked bar plot
Version: 0.0.0.9000
Authors@R: 
    person(given="Noah", family="Irwin",  email = "noah.irwin@selu.edu", role = c("aut", "cre"))
Description: What the package does (one paragraph).
License: MIT
Encoding: UTF-8
Roxygen: list(markdown = TRUE)
RoxygenNote: 7.3.2

```{r}
##use_roxygen_md()

#then open one of the R fucntions in the R folder
#Then wehn you open it click the very beginning of line one. Then click the magic wand at the top and choose insert roxygen skeleton
```
the R fucntion should look liek this when fileld out

#' Creates a stacked bar chart when given columns and a dataframe
#'
#' @param data The dataframe you want to use
#' @param column1 The name of the first column of data applicable found in your dataframe
#' @param column2 The name of the second column of data applicable found in your dataframe
#' @param column3 The name of the third column of data applicable found in your dataframe
#' @param graphname The name you want to give to the graph you are created by this function
#'
#' @return plot  A ggplot object of the posterior trace
#' @export
stack_function<-function(data,column1,column2,column3,graphname){
  graph_data<-data %>%
    select({{column1}},{{column2}},{{column3}}) %>%
    gather(key=sgls, value=mm) %>%
    mutate(rounded = round(mm)) %>%
    group_by(rounded, sgls) %>%
    summarize(count=n())
  plot<-ggplot(data=graph_data,mapping=aes(x=rounded,y=count,fill=sgls))+geom_col(position = position_stack())+xlab("SGL's (mm)")+ylab("Frequency")
  ggsave(graphname, width = 160, height = 80, units = "mm",dpi = 300)
  return(graph_data)
}
```{r}
#document()  creates a manual for every fucntion that has documentation
#use_package("name of the package you used for a specific function")
#use_pipe()
```
open man file in the project section

do not edit the maunal file change the fucntion section and then reload to make changes

```{r}
##install() then choose 3
```

building a tutorital

```{r}
##use_vignette("purrfect","Project three functions")
```

check()
this fucntion checks for a variety of errors in your R code functions

installing packages from github

```{R}
devtools::install_github("noahsirwin/Project_three")

```
