---
title: "Thesis Functions and Variables"
output: html_document
date: "2024-10-01"
---
install.packages("readxl")
library(readxl)

##Collection 1 sheets
athens_sgl_1<-readxl::read_excel("data/First_collection_sgls.xlsx", na = c("NA", ""),sheet="athens1")
cochran_sgl_1<-readxl::read_excel("data/First_collection_sgls.xlsx",na = c("NA", ""),sheet="cochran1")
auburn_sgl_1<-readxl::read_excel("data/First_collection_sgls.xlsx",na = c("NA", ""),sheet="auburn1")
tusk_sgl_1<-readxl::read_excel("data/First_collection_sgls.xlsx",na = c("NA", ""),sheet="tuskegee1")
pasc_sgl_1<-readxl::read_excel("data/First_collection_sgls.xlsx",na = c("NA", ""),sheet="pascagoula1")
broad_sgl_1<-readxl::read_excel("data/First_collection_sgls.xlsx",na = c("NA", ""),sheet="broadriver1")
bogue_sgl_1<-readxl::read_excel("data/First_collection_sgls.xlsx",na = c("NA", ""),sheet="boguechitto1")

##Collection 2 sheets
athens_sgl_2<-readxl::read_excel("data/Second_collection_sgls.xlsx",na = c("NA", ""),sheet="athens2")
cochran_sgl_2<-readxl::read_excel("data/Second_collection_sgls.xlsx",na = c("NA", ""),sheet="cochran2")
auburn_sgl_2<-readxl::read_excel("data/Second_collection_sgls.xlsx",na = c("NA", ""),sheet="auburn2")
tusk_sgl_2<-readxl::read_excel("data/Second_collection_sgls.xlsx",na = c("NA", ""),sheet="tuskegee2")
pasc_sgl_2<-readxl::read_excel("data/Second_collection_sgls.xlsx",na = c("NA", ""),sheet="pascagoula2")
broad_sgl_2<-readxl::read_excel("data/Second_collection_sgls.xlsx",na = c("NA", ""),sheet="broadriver2")
bogue_sgl_2<-readxl::read_excel("data/Second_collection_sgls.xlsx",na = c("NA", ""),sheet="boguechitto2")

##Collection 3 sheets
athens_sgl_3<-readxl::read_excel("data/Third_collection_sgls.xlsx",na = c("NA", ""),sheet="athens3")
cochran_sgl_3<-readxl::read_excel("data/Third_collection_sgls.xlsx",na = c("NA", ""),sheet="cochran3")
auburn_sgl_3<-readxl::read_excel("data/Third_collection_sgls.xlsx",na = c("NA", ""),sheet="auburn3")
tusk_sgl_3<-readxl::read_excel("data/Third_collection_sgls.xlsx",na = c("NA", ""),sheet="tuskegee3")
pasc_sgl_3<-readxl::read_excel("data/Third_collection_sgls.xlsx",na = c("NA", ""),sheet="pascagoula3")
broad_sgl_3<-readxl::read_excel("data/Third_collection_sgls.xlsx",na = c("NA", ""),sheet="broadriver3")
bogue_sgl_3<-readxl::read_excel("data/Third_collection_sgls.xlsx",na = c("NA", ""),sheet="boguechitto3")

##Collection 4 sheets
athens_sgl_4<-readxl::read_excel("data/Fourth_collection_sgls.xlsx",na = c("NA", ""),sheet="athens4")
cochran_sgl_4<-readxl::read_excel("data/Fourth_collection_sgls.xlsx",na = c("NA", ""),sheet="cochran4")
auburn_sgl_4<-readxl::read_excel("data/Fourth_collection_sgls.xlsx",na = c("NA", ""),sheet="aubrun4")
tusk_sgl_4<-readxl::read_excel("data/Fourth_collection_sgls.xlsx",na = c("NA", ""),sheet="tuskegee4")
pasc_sgl_4<-readxl::read_excel("data/Fourth_collection_sgls.xlsx",na = c("NA", ""),sheet="pascagoula4")
broad_sgl_4<-readxl::read_excel("data/Fourth_collection_sgls.xlsx",na = c("NA", ""),sheet="broadriver4")
bogue_sgl_4<-readxl::read_excel("data/Fourth_collection_sgls.xlsx",na = c("NA", ""),sheet="boguechitto4")


##Journal Stream Data 1 sheets
athens_jour_1<-readxl::read_excel("data/journal_data1.xlsx",na = c("NA", ""),sheet="athens_jour_1")
cochran_jour_1<-readxl::read_excel("data/journal_data1.xlsx",na = c("NA", ""),sheet="cochran_jour_1")
auburn_jour_1<-readxl::read_excel("data/journal_data1.xlsx",na = c("NA", ""),sheet="auburn_jour_1")
tusk_jour_1<-readxl::read_excel("data/journal_data1.xlsx",na = c("NA", ""),sheet="tuskegee_jour_1")
pasc_jour_1<-readxl::read_excel("data/journal_data1.xlsx",na = c("NA", ""),sheet="pascagoula_jour_1")
broad_jour_1<-readxl::read_excel("data/journal_data1.xlsx",na = c("NA", ""),sheet="broad_jour_1")
bogue_jour_1<-readxl::read_excel("data/journal_data1.xlsx",na = c("NA", ""),sheet="bouge_jour_1")

##Journal Stream Data 2 sheets
athens_jour_2<-readxl::read_excel("data/journal_data2.xlsx",na = c("NA", ""),sheet="athens_jour_2")
cochran_jour_2<-readxl::read_excel("data/journal_data2.xlsx",na = c("NA", ""),sheet="cochran_jour_2")
auburn_jour_2<-readxl::read_excel("data/journal_data2.xlsx",na = c("NA", ""),sheet="auburn_jour_2")
tusk_jour_2<-readxl::read_excel("data/journal_data2.xlsx",na = c("NA", ""),sheet="tuskegee_jour_2")
pasc_jour_2<-readxl::read_excel("data/journal_data2.xlsx",na = c("NA", ""),sheet="pascagoula_jour_2")
broad_jour_2<-readxl::read_excel("data/journal_data2.xlsx",na = c("NA", ""),sheet="broad_jour_2")
bogue_jour_2<-readxl::read_excel("data/journal_data2.xlsx",na = c("NA", ""),sheet="bouge_jour_2")

##Journal Stream Data 3 sheets
athens_jour_3<-readxl::read_excel("data/journal_data3.xlsx",na = c("NA", ""),sheet="athens_jour3")
cochran_jour_3<-readxl::read_excel("data/journal_data3.xlsx",na = c("NA", ""),sheet="cochran_jour_3")
auburn_jour_3<-readxl::read_excel("data/journal_data3.xlsx",na = c("NA", ""),sheet="aubrun_jour_3")
tusk_jour_3<-readxl::read_excel("data/journal_data3.xlsx",na = c("NA", ""),sheet="tuskegee_jour_3")
pasc_jour_3<-readxl::read_excel("data/journal_data3.xlsx",na = c("NA", ""),sheet="pascagoula_jour_3")
broad_jour_3<-readxl::read_excel("data/journal_data3.xlsx",na = c("NA", ""),sheet="broad_jour_3")
bogue_jour_3<-readxl::read_excel("data/journal_data3.xlsx",na = c("NA", ""),sheet="bogue_jour_3")

##Journal Stream Data 4 individual sheets
athens_jour_4<-readxl::read_excel("data/journal_data4.xlsx",na = c("NA", ""),sheet="athens_jour_4")
cochran_jour_4<-readxl::read_excel("data/journal_data4.xlsx",na = c("NA", ""),sheet="cochran_jour_4")
auburn_jour_4<-readxl::read_excel("data/journal_data4.xlsx",na = c("NA", ""),sheet="auburn_jour_4")
tusk_jour_4<-readxl::read_excel("data/journal_data4.xlsx",na = c("NA", ""),sheet="tuskegee_jour_4")
pasc_jour_4<-readxl::read_excel("data/journal_data4.xlsx",na = c("NA", ""),sheet="pascagoula_jour_4")
broad_jour_4<-readxl::read_excel("data/journal_data4.xlsx",na = c("NA", ""),sheet="broad_jour_4")
bogue_jour_4<-readxl::read_excel("data/journal_data4.xlsx",na = c("NA", ""),sheet="bogue_jour_4")

##Journal comparative sheets
wb_sheets <- readxl::excel_sheets("data/journal_data1.xlsx")
data_frame_jour1 = lapply(setNames(wb_sheets, wb_sheets),function(x) read_excel("data/journal_data1_amw.xlsx", sheet=x)) 
df_jour1 = bind_rows(data_frame_jour1, .id="Sheet")                                                          
df_jour1

wb_sheets2 <- readxl::excel_sheets("data/journal_data2.xlsx") 
data_frame_jour2= lapply(setNames(wb_sheets2, wb_sheets2),function(x) read_excel("data/journal_data2.xlsx", sheet=x)) 
df_jour2 = bind_rows(data_frame_jour2, .id="Sheet")                                                          
df_jour2

wb_sheets3 <- readxl::excel_sheets("data/journal_data3.xlsx") 
data_frame_jour3= lapply(setNames(wb_sheets3, wb_sheets3),function(x) read_excel("data/journal_data3.xlsx", sheet=x)) 
df_jour3 = bind_rows(data_frame_jour3, .id="Sheet")                                                          
df_jour3

wb_sheets4 <- readxl::excel_sheets("data/journal_data4.xlsx") 
data_frame_jour4= lapply(setNames(wb_sheets4, wb_sheets4),function(x) read_excel("data/journal_data4.xlsx", sheet=x)) 
df_jour4 = bind_rows(data_frame_jour4, .id="Sheet")                                                          
df_jour4


##Ecoregion
df_jour1_coastal<-df_jour1[c(1:3),]
df_jour1_piedmont<-df_jour1[c(4:7),]

df_jour2_coastal<-df_jour2[c(1:3),]
df_jour2_piedmont<-df_jour2[c(4:8),]

df_jour3_coastal<-df_jour3[c(1:3),]
df_jour3_piedmont<-df_jour3[c(4:8),]

df_jour4_coastal<-df_jour4[c(1:3),]
df_jour4_piedmont<-df_jour4[c(4:8),]


##Combined Ecoregions

df_coastal_plain_combined=bind_rows(list(df_jour1_coastal,df_jour2_coastal,df_jour3_coastal,df_jour4_coastal), .id= "Coastal")
df_coastal_plain_ecoregion<-df_coastal_plain_combined %>% 
mutate(Coastal="coastal")

df_piedmont_combined=bind_rows(list(df_jour1_piedmont,df_jour2_piedmont,df_jour3_piedmont,df_jour4_piedmont), .id = "Piedmont")
df_piedmont_combined_ecoregion<-df_piedmont_combined %>% 
 mutate(Piedmont="piedmont")

###Combined whole journal data
df_jour_whole_combined=bind_rows(df_piedmont_combined_ecoregion,df_coastal_plain_ecoregion)
##combined with a new ecoregion column for comparisons
master_df_jour_combined<-df_jour_whole_combined %>%
unite(col=Ecoregion,Piedmont,Coastal)




###Functions

##Working
THE_function<-function(filepath,columns){
dataframe<-read_excel(filepath,na = c("NA", ""))
 data<-dataframe %>% 
 pull(columns)
avg<-mean(data)
return(avg)
}
filepath_sgl1<-"thesis_project/First_collection_sgls.xlsx"

##Working
Thesis_avg_function<-function(data,columns){
    data<-data%>% 
        na.omit(dataframe) %>% 
        pull(columns)
    avg<-mean(data)
    return(avg)
}
# Thesis_avg_function Input
Thesis_avg_function(athens_sgl_1,"larva_sgls")


##(Broken) (select in stead of pull)
 Frequency_function<-function(data,column1,column2,column3){
     col1<-data %>% 
          na.omit(dataframe) %>% 
           select({{column1}}) %>% 
           count()
 
 #      N_total<-col_1+col_2+col3
 #       if(col_2>0 | col_3>0){
 #          return(N_total)
 #}else{    
 message(paste("Column2 or Column3 is empty.",  col1))
 }

###Frequency function first rendition
Frequency_function<-function(data,column1,column2,column3){
  
  col_1<-data %>% 
    select({{column1}}) %>% 
    na.omit() %>% 
    count()
  col_2<-data %>%
    select({{column2}}) %>%
    na.omit() %>% 
    count()
  col_3<-data %>%
    select({{column3}}) %>%
    na.omit() %>%
    count()
  N_total<-col_1+col_2+col_3
  if(col_2>0 | col_3>0){
    return(N_total)
    }else{
      message(paste("Column2 or Column3 is empty.", col_1))
      }
}
##Input
Frequency_function(data=athens_sgl_1,column1=larva_sgls,column2=meta_sgls,column3=adult_rounded_sgls)

#This isnt working as I would like. There are 125 rows in column1 but there is no data in column2 and column3 so I want it to be 125 + 0 + 0 but rigth now its counting 125 + 1 + 1


stack_function<-function(data,column1,column2,column3,graphname){
     graph_data<-data %>%
         select({{column1}},{{column2}},{{column3}}) %>%
         gather(key=sgls, value=mm) %>% 
         mutate(rounded = round(mm)) %>% 
         group_by(rounded, sgls) %>% 
         summarize(count=n())
     plot<-ggplot(data=graph_data,mapping=aes(x=rounded,y=count,fill=sgls))+geom_col(position = position_stack())+xlab("SGL's (mm)")+ylab("Frequency")
     ggsave(graphname, width = 160, height = 80, units = "mm",dpi = 300)
     return(graph_data)
 }
 ##Input
stack_function(data=athens_sgl_4,column1=larva_sgls,column2=meta_sgls,column3=adult_rounded_sgls,graphname="data/plots/athens_sgl_4_plot.png")


#Binding and pairing like sheets into a single data frame
library(vctrs)
wb_sheets <- readxl::excel_sheets("data/"".xlsx") 
"data_frame" = lapply(setNames(wb_sheets, wb_sheets),  
                     function(x) read_excel("data/"".xlsx", sheet=x)) 
df_"" = bind_rows("data_frame", .id="Sheet")                                                   
df_""


##Mapping Function
```{r}
mapping_function<-function(data,column1,column2,column3){
  mapping_data<-data %>% 
    select({{column1}},{{column2}},{{column3}}) %>%
    group_by({{column3}}) %>% 
    na.omit() %>% 
    mutate(Actual_lon=as.integer({{column1}})) %>% 
    mutate(Actual_lat=as.integer({{column2}}))
  
  map<-leaflet::leaflet(mapping_data) %>%
      addTiles() %>% 
      addMarkers(~Actual_lon,~Actual_lat,popup=mapping_data$column3)
  return(map)
  
}
    
    #map<-leaflet::leaflet(mapping_data) %>%
      #addTiles() %>% 
      #addMarkers(~{{column1}},~{{column2}},popup=mapping_data$column3)
  #return(function_map)

  #mapcoords<-mapping_data %>% 
    #st_as_sf(coords=c("Actual_lon","Actual_lat"),crs=4326)
  
  #function_map<-mapview(mapcoords,zcol="Locality",cex="Locality",layer.name="Locality",label="Locality") %>%  
  #addStaticLabels(label=mapping_data$Locality,noHide=TRUE,direction="auto",offset=c(0,10),textOnly=TRUE,textsize="12px")   
  
  

```
##Input
mapping_function(data=Thesis_Data_Table_for_R_and_Figures,column1=Actual_lon,column2=Actual_lat,column3=Locality)


##Mapping toolshed
north_coords<-shan_set %>% 
  st_as_sf(coords=c("Actual.Lon","Actual.Lat"),crs=4326)

camMap_labeled<-mapview(north_coords,zcol="qD",cex="qD",layer.name="qD",label="Deployment.Name") %>% 
  addStaticLabels(label=shan_set$Deployment.Name,noHide=TRUE,direction="auto",offset=c(0,10),textOnly=TRUE,textsize="12px")
camMap_labeled  

camMap<-mapview(deployCount_sf,zcol="Subproject",cex="new_count",layer.name="Habitat.Type") ##zcol colors by "subproject"


##Linear_function
```{r}
Linear_function<-function(data,x=column1,y=column2){
     linear_dataframe<-data %>% 
         select({{column1}},{{column2}}) %>% 
         na.omit() %>%
         model_fit<-lm({{column2}}~{{column1}},) %>% 
             summary<-summary(model_fit) %>% 
             return(summary)
 }

```

##Input linear_function
Linear_function(Thesis_Data_Table_for_R_and_Figures,x=Avg_temp,y=Avg_sample_sgl)

###Input Linear Plot function
```{r}
Linear_plot_function<-function(data,x=column1,y=column2,xlab=text1,ylab=text2,title=text3,label=text4){
line_plot<-ggplot(data,aes(x={{column1}},y={{column2}}))+geom_point()+geom_smooth(method="lm",color="black",size=0.5,fill="gray")+ labs(x="{{text1}}",y="{{text2}}", title="{{text3}}")+annotate("text",x=20,y=28,label="{{text4}}",parse=T)+theme_bw()
  return(line_plot)
}
```
###Input Linear Plot function
Linear_plot_function(Thesis_Data_Table_for_R_and_Figures,x=Avg_temp,y=Avg_sample_sgl,xlab="Avg temperature(celcius)",ylab="Avg Size SGL(mm)",title="Linear regression to predict salamander size",label="R^2==0.966")



