---
title: "Homework One"
output: html_document
---

## Homework One

### Due Aug. 30th at 5pm. Simply write your answers in and save in the homeworks folder in your Rstudio.

This is going to be a low-stakes effort to get you comfortable with RStudio and RMarkdown. First, you're going to download the RMarkdown Document for this homework. Copy and paste this command into the console of your RStudio instance.

```{r}
#download.file(url = "https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2024/main/vignettes/homework_1.Rmd", destfile = "/cloud/project/homeworks/Homework_1.Rmd")
```

Can you find where the file downloaded to? Look at the help entry for `download.file` if you aren't sure.

Open the file and answer these two questions: 

1. (1pt) What does `download.file` do? 

```
Downloads a file to R-studio. If it ends in ".rmd" it is displayed in the top left "source" pane as a R-markdown page.
```
2. (1pt) Where did your Homework_1.Rmd file download to? 

```
My "project" folder. More specifically to a folder found in my projects folder titled "homeworks"

```

Next, we're going to try to execute some code in an RMarkdown document. Press the green triangle.

```{r}
animals <- c("cat", "rat", "dog") #assigns the variable animals as a vector. The vector contains 3 characters ("cat", "rat", "dog).
weights <- c(15, 1, 120)
animals[c(1,3)] #Displays the 1st and 3rd characters of the animal vector (cat and dog).
```

What just happened? Annotate the above code with two comments (each worth one point) explaining the code.

We can also embed plots in RMarkdown. The below code should embed a barplot. Not very useful, but let me know if you don't see one show up in your file.

```{r}
barplot(names.arg=animals, height=weights)
```

Did the plot show up OK?

```
Yes

```

Given the above code, what do you think the X and Y axes represent on this graph? 

```
X axis is weights and Y axis is animals

```


## Getting to know you all - Grad Students

For the remainder of the assignment, I need to know four things.

1. What type of data will you be working with for your thesis or research project? For example: matrices of phylogenetic characters, spreadsheets of ecological observations, shapefiles from 3D scanning, etc. If you aren't actively doing research, what kinds of scientific questions do you find interesting?

#I need to run mixdist on some  SVL_lengths. The best fit model will tell me how many age classes best fit my samples. I also need to crunch a bunch of abiotic stream data and see if they have any relevance to the variablitiy in number of age classes. I was thinking a two-way ANOVA for this, but am unsure.

#AMW: Hmmm, yeah, two way ANOVA. Or maybe a non-linear model? We can see when we get to that part of class. 

2. What are you hoping to get out of class?

#I want to understand R-studio enough that I can analyize my own data and potentially use my understanding of R in the job field. I'm just trying to become  indispensable at my next job.

3. Are there topics that aren't on the schedule that you'd be interested to see? 
#Maybe Bayesian Models. I don't think I understand enough of R's capabilties to know what is possible in R-studio.
#AMW: I don't think we'll get to that, exactly, but we might do some simulation.

4. Do you know in advance of any absences I should be aware of? 
I will not be in class Tuesday 2024_09_03 as I will be finishing up my final sampling trip.