---
title: "Project Three Functions Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Project three functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introducing the Purrfect package
+ This package was created as the final project of an R course. 
+ It's purpose is to aid me in the analysis and plotting of my thesis data. 
+ This includes generating averages, frequencys, linear regressions, and plots. 
+ This package works with any data frame that has been read into R. 

## Installation Instructions

```{r setup}
devtools::install_github("noahsirwin/Project_three")
library(purrfect)
```

## Where to get the data
```{r}
library(readxl)
library(googlesheets4)
library(readxl)
gs4_deauth()
Thesis_Data_Table_for_R_and_Figures<-readxl::read_excel("../data/Thesis Data Table for R and Figures.xlsx")
```
## Using the Thesis_avg_function
```{r eval=FALSE, warning=FALSE}
Thesis_avg_function<-function(data,columns){
  data<-data%>%
    na.omit(dataframe) %>%
    pull(columns)
  avg<-mean(data)
  return(avg)
}
```

## Using the stack_function
```{r eval=FALSE, warning=FALSE}
stack_function<-function(data,column1,column2,column3,graphname){
  graph_data<-data %>%
    dplyr::select({{column1}},{{column2}},{{column3}}) %>%
    tidyr::gather(key=sgls, value=mm) %>%
    dplyr::mutate(rounded = round(mm)) %>%
    dplyr::group_by(rounded, sgls) %>%
    dplyr::summarize(count=n())
  plot<-ggplot2::ggplot(data=graph_data,mapping=aes(x=rounded,y=count,fill=sgls))+ggplot2::geom_col(position = position_stack())+xlab("SGL's (mm)")+ylab("Frequency")
  ggplot2::ggsave(graphname, width = 160, height = 80, units = "mm",dpi = 300)
  return(graph_data)
}
```

## Using the Frequency_function 
```{r eval=FALSE, warning=FALSE}
Frequency_function<-function(data,column1,column2,column3){
  col_1<-data %>%
    dplyr::select({{column1}}) %>%
    na.omit() %>%
    count()
  col_2<-data %>%
    dplyr::select({{column2}}) %>%
    na.omit() %>%
    count()
  col_3<-data %>%
    dplyr::select({{column3}}) %>%
    na.omit() %>%
    count()
  N_total<-col_1+col_2+col_3
  if(col_2>0 | col_3>0){
    return(N_total)
  }else{
    message(paste("Column2 or Column3 is empty.", col_1))
  }
}
```

## Using Function 4
```{r eval=FALSE, warning=FALSE}
```

## Using Function 5
```{r eval=FALSE, warning=FALSE}
```
