---
title: "NOTES10-29-24"
author: "Noah Irwin"
date: "2024-10-29"
output: html_document
---

install.packages("remotes")
remotes::install_github("r-spatial/mapview")

If you have a package that will not load properaly when installing you might have a problem where a busted directory was created and all future packages are trying to be downloaded to them. Mine looked like  this

```
Installing package into ‘/cloud/lib/x86_64-pc-linux-gnu-library/4.4’
(as ‘lib’ is unspecified)
ERROR: failed to lock directory ‘/cloud/lib/x86_64-pc-linux-gnu-library/4.4’ for modifying
Try removing ‘/cloud/lib/x86_64-pc-linux-gnu-library/4.4/00LOCK-mapview’
Warning message:
In i.p(...) :
  installation of package ‘/tmp/Rtmpc3XAF7/file14a1fb407e8/mapview_2.11.2.9000.tar.gz’ had non-zero exit status

```
copy the section of code that starts with    Try removing '---------'
to address this problem go to the terminal tab of the console pane and paste in the section you copied:
rm -r /cloud/lib/x86_64-pc-linux-gnu-library/4.4/00LOCK-mapview

```{r}
using<-function(...) {
    libs<-unlist(list(...))
    req<-unlist(lapply(libs,require,character.only=TRUE))
    need<-libs[req==FALSE]
    if(length(need)>0){ 
        install.packages(need)
        lapply(need,require,character.only=TRUE)
    }
}
packagesInThisExercise <- c("tidyverse", "googlesheets4", "iNEXT", "fields", "sf", "mapview")
using(packagesInThisExercise)


```

The Data Sheet
```{r}
library(tidyverse)
library(googlesheets4)
gs4_deauth()
cam <- read_sheet("https://docs.google.com/spreadsheets/d/1esrbC-QqoufjlDHXtIYtDcsXovC_x6o6Z4JfBPKvg3s/edit?usp=sharing")
```

```{r}
##Making names
names(cam)<-make.names(names(cam),unique=TRUE)
names(cam)

cam$Species.Name<-str_replace_all(cam$Species.Name," ","_") ##replaces spaces with underscores 
cam$Species.Name
```


Filtering out extra data
```{r}
speciesCount<-cam %>%
  select(Species.Name,Count,Subproject) %>%
  filter(Species.Name !="Calibration.Photos",
        Species.Name !="Camera.Misfire",
        Species.Name !="No.Animal",
        Species.Name !="Camera.Trapper",
        Species.Name !="Animal.Not.on.List",
        Species.Name !="Vehicle",
        Species.Name !="Homo.Sapiens") %>% 
        group_by(Species.Name,Subproject) %>% 
          summarize(totalCount=sum(Count)) %>% 
          spread(Subproject,totalCount)

speciesCount
```

Species Abundance
```{r}
abundance<-cam %>% 
  group_by(Subproject) %>% 
  summarize(SpeciesAbuncance=length(unique(Species.Name)))
abundance
```

install.packages("iNext)

```{r}
speciesCount[is.na(speciesCount)]=0
camDiv<-as.data.frame(speciesCount) ##Beacue iNext needs base R dataframes
camDiv
```

```{r}
rownames(camDiv)<-camDiv$Species.Name
camDiv<-camDiv[2:6]
camDiv
```

Species Richness
```{r}
##library(iNEXT)
richness<-iNEXT(camDiv,q=0,datatype="abundance") ##calcualtes species richness q=0
richness
```

Creating a species richness plot
```{r}
ggiNEXT(richness)+theme(legend.position="right")+theme_classic()

#species density is richness on the y axis
```

Shannon diversity index  (Species Mixing)
 calculating the proportion of the total individual that there were specific species 
```{r}
diversity_Shan <-iNEXT(camDiv,datatype = "abundance",q=1) ## q=1 is the shannon index
ggiNEXT(diversity_Shan)+theme(legend.position = "right")+theme_classic()
```

Simpsons diversity index (Rarity)
number organisms that belong to each species normalized by the total number of organisms
0 is a lower probability that you will draw the same animal randomly 

```{r}
diversity_Simps <-iNEXT(camDiv,datatype = "abundance",q=2) ## q=2 is the Simpsons index
ggiNEXT(diversity_Simps)+theme(legend.position = "right")+theme_classic()
```

```{r}
allDiversity<-iNEXT(camDiv,q=c(0,1,2),datatype="abundance")
ggiNEXT(allDiversity,facet.var="Order.q")+theme(legend.position="right")+theme(axis.text.x = element_text(colour="black",size=12,angle=45,hjust=.5,vjust=.05))+theme_classic ()
```
You should also consider how close some communties are to eachother. You migth be picking up transiant species
ggpolt(cam,aes(x=Actual.Lon,y= Actual.Lat

```{r}
deployCount<-cam %>% 
select(Species.Name,Count,Deployment.ID,Actual.Lon,Actual.Lat,Subproject) %>% 
group_by(Deployment.ID,Actual.Lon,Actual.Lat,Subproject) %>% 
  summarize(new_count=sum(Count))

deployCount

ggplot(deployCount)+geom_point(aes(x=Actual.Lon,y=Actual.Lat,size=new_count,color=Subproject),alpha=0.5)+theme_bw()
```

Maps Trnasfor,ing the corrdiantes the the proprityory form it needs
```{r}
##library(sf)
deployCount_sf<-deployCount %>% 
  st_as_sf(coords=c("Actual.Lon","Actual.Lat"),crs=4326)
deployCount_sf
```

```{r}
##library(mapview)
camMap<-mapview(deployCount_sf,zcol="Subproject",cex="new_count",layer.name="Habitat.Type") ##zcol colors by "subproject"
camMap
```
