---
title: "NotesActual11-5-2024"
author: "Noah Irwin"
date: "2024-11-05"
output: html_document
---
```{r}
ggplot(cam,aes(x=Actual.Lon ,y=Actual.Lat,color=Subproject))+geom_point()+theme_bw()
```
symmetric difference the number of unique groupings in a tree when comparing two trees
summary statistic-some number that summarizes an individual or population

11x        11x    4

1x11       1x11   6

11x        11x    4
ordination display things spacially based on their summary statistics


Bray curtis distance

2->3
minimum 
crow 0
sparrow 6
tern 0
robin 4
blue Heron 0
cardinal 0
woodpeaker 0
thrush 0
minimum = 10  Cjk
total # site 2 = 33  Sj
total # site 3 = 16  Sk
BCjk=1- (2*Cjk/Sj+Sk) = 1-.408= .60

1->2
BC=1.0

the scale is 0-1    0=the exact same    1=completely different
1->3
minimum
red wing 0
crow 4
red tail 0
green hrn 0
bl hrn 1
sparrow 0
tern 0
robin 0

minimum = 5 CjK
total 1 =19
total 3= 16
BC= 1-.2857 ==.714

install.package("vegan")
library(vegan)
```{r}
speciesCount<-cam %>%
  group_by(Common.Name,Subproject) %>% 
  summarize(total=sum(Count)) %>% 
  spread(Common.Name,total,fill=0)
speciesCount

```
Vegan only works if it is in matrix format
```{r}
speciesMatrix<-data.matrix(speciesCount,rownames.force=unique(cam$Subproject))
speciesMatrix
```

```{r}
cam_trap_NMDS<-metaMDS(speciesMatrix)
cam_trap_NMDS
```
its alright if stress is nearly zero this just means that the points adn their distances from each other are easily plotted 

Plotting NMDS
```{r}
ordiplot(cam_trap_NMDS,type="n") #RUN THIS BEFORE ON ITS OWN
orditorp(cam_trap_NMDS,display="sites",cex=1.25)  #sites comes from an object
```

pulling data out to put on ggplot
```{r}
data.scores=as.data.frame(scores(cam_trap_NMDS)$sites)
data.scores
#sites are in the order they appear in our original speciescount data base
```
plotting out now with ggplot

```{r}
names<-data.scores %>% 
  mutate(Location=speciesCount$Subproject)
names ##This reaplys the localtity names to our varaibales
```

```{r}
ggplot(data=names, aes(x=NMDS1 ,y=NMDS2))+geom_point(size=4,aes(colour=Location))
#you can plot this without an axis
```

on your own
library(googlesheets4)
```{r}
gs4_deauth()##dont want to eb asked to sign into google when accessing publicllay available data
nop <- read_sheet("https://docs.google.com/spreadsheets/d/18yhZcaiYt6llg84eZYrNEewuLP3xaEyxYQQN__RQ3_4/edit?usp=sharing")
view(nop)
```

Who had the most Deer

preparing data first
```{r}
names(nop)<-make.names(names(nop),unique=TRUE)
names(nop)
nop$Common.Name<-str_replace_all(nop$Common.Name," ","_") ##replaces spaces with underscores 
nop$Species.Name<-str_replace_all(nop$Species.Name," ",".")
nop$Subproject<-str_replace_all(nop$Subproject," ",".")
nop$Subproject<-str_replace_all(nop$Subproject,"/",".")
nop$Common.Name
```


check to see if there is variation
```{r}
num_deer<-nop %>% 
  filter((Common.Name=="White-tailed_Deer"))%>% 
  group_by(Deployment.Name)
ggplot(num_deer,aes(x=Deployment.Name))+geom_bar()
##Only five of the sites show up in the bar plot so we know there is a discrepancy in deer number per site
```
library
```{r}
deer_survey<-nop %>%
  filter(Common.Name =="White-tailed_Deer") %>%  ## "==" find only  "!" leave out
  group_by(Actual.Lat,Actual.Lon,Deployment.Name) %>% 
  summarise(count=n())
deer_survey  
ggplot(deer_survey, aes(x=Actual.Lon, y=Actual.Lat,color=count,size=count)) + geom_point()
```
turning Deer survey into a usable base R dataframe
```{r}
speciesCount_dEER<-nop %>% 
  select(Common.Name,Count,Deployment.Name) %>% 
  group_by(Common.Name,Deployment.Name) %>% 
  summarise(totalCount= sum(Count)) %>% 
  spread(Deployment.Name,totalCount)


speciesCount_dEER[is.na(speciesCount_dEER)]=0
nopdiv<-as.data.frame(speciesCount_dEER)
rownames(nopdiv)<-nopdiv$Common.Name
nopdiv<-nopdiv[2:11] ##As there was a total or 10 camera traps so we need to do 11. This is because computers start counting at 1   camera 2 is the first camera trap column and we go to the last column wtih camera trap information 9
```
```{r}
nop_iNEXT<-iNEXT(nopdiv,q=1,datatype = "abundance")
nop_iNEXT
```
mapping
```{r}
ggiNEXT(nop_iNEXT)+theme(legend.position = "right")+theme_classic()
```

Place the shannon index unit into its own data frame
```{r}

df<-as.data.frame(nop_iNEXT$iNextEst$size_based)
df
###We only want the methods sections titled "observed"
shannon_df_clean<-df %>% 
 filter(Method=="Observed") %>%
  select(Assemblage,qD) %>% 
  rename("Deployment.Name"= Assemblage)

shannon_df_clean
```
In order to plot this we need a few things Deployment.Name,Actual.Lat,Actual.lon as well as our shannon index information so we must join two dataframes
```{r}
shan_set<-full_join(shannon_df_clean,nop,by="Deployment.Name")
shan_set
```

Plotting the information
```{r}
ggplot(shan_set,aes(x=Actual.Lon,y=Actual.Lat,color=qD))+geom_point(aes(alpha=.5,color=))+theme_bw()

#You need a package 
#install.packages("ggrepel")
#library(ggrepel) this allows you to label individual points with the argument label= "" within ggplot()

#labeling these points by Deployment.Name
#ex.
#ggplot(shan_set,aes(x=Actual.Lon,y=Actual.Lat,color=qD,label=Deployment.Name))+geom_point(aes(alpha=.5,color=))+theme_bw()

```
mapping
```{r}
north_coords<-shan_set %>% 
  st_as_sf(coords=c("Actual.Lon","Actual.Lat"),crs=4326)

camMap<-mapview(north_coords,zcol="qD",cex="qD",layer.name="qD",label="Deployment.Name")
camMap

##applies labels to the map by using a legend
#we need another package
#install.packages("leafem")
#library(leafem)
camMap_labeled<-mapview(north_coords,zcol="qD",cex="qD",layer.name="qD",label="Deployment.Name") %>% 
  addStaticLabels(label=shan_set$Deployment.Name,noHide=TRUE,direction="auto",offset=c(0,10),textOnly=TRUE,textsize="12px")
camMap_labeled
```



