---
title: "ClassNotes"
author: "Noah Irwin"
date: "2024-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here are my notes from class 08-22-2024
"<-" is the assignment variable
"something" assigns the variable to something
if you see + thta means that R is asking for more info
R studio doesnt want names to start with numbers
Cntrl+L clears the consule pane
A vector is a one deminsional grouping of objects lists and vectors may be used interchangable

to get to another lien when in {} hit shift + enter
for (weight in weight_g){
print(weight/1000)
}
For each weight divide the weight in the "For loop" by 1000

## Variables
A variable is soemthing R holds in Memory. You can makea variable like so:
```{r}
weight_kg <- 50
```

## Vectors
Vectors are list like grouping of objects. A vector looks like this:
```{r}
weight_g <- c(50,60,65,70,80,90)
weight_g
```

## Varibale Reassignment
```{r}
mass <- 47.5
mass <- mass * 2.2 #I am reassigning mass to be in lbs
```

## Class
type class (animls) or class (weight_g) to ensure that you have them correctly organized as either numerical or character

> weight
[1] 90
> weight_g
[1] 50 60 65 70 80 90
> weight_g <- c(weight_g, 90)
> weight_g
[1] 50 60 65 70 80 90 90
 can be used to add a number to a previous list
 
 ## And/Or 
  weight_g
[1] 30 50 60 65 70 90 90
> weight_g[weight_g >50| weight_g >70]
[1] 60 65 70 90 90
> weight_g[weight_g <50| weight_g >70]
[1] 30 90 90
> ##| is the "or" symbol    & is the "and" symbol

## TO REMOVE NAs when you need an average
heights <- c(2,4,4,NA,6)
> mean(heights)
[1] NA
> mean(heights, na.rm = TRUE)

## Entering Surveys as a Vector
head(surveys[, "species_id"])

## Entering Surveys as a data frame
head(surveys["species_id"])

## Challange 
> surveys_200 <- surveys [200,]
> surveys_200
    record_id month day year plot_id species_id sex hindfoot_length
200     35212    12   7 2002       2         NL   M              33
    weight   genus  species   taxa plot_type
200    248 Neotoma albigula Rodent   Control
> nrows(surveys)
Error in nrows(surveys) : could not find function "nrows"
> nrow(surveys)
[1] 34786
> surveys[34786]
Error in `[.data.frame`(surveys, 34786) : undefined columns selected
> surveys[34786,]
      record_id month day year plot_id species_id sex hindfoot_length
34786     30986     7   1 2000       7         PX                  NA
      weight       genus species   taxa        plot_type
34786     NA Chaetodipus     sp. Rodent Rodent Exclosure

##or
surveys_last <- surveys[nrow(surveys),]
> surveys_last
      record_id month day year plot_id species_id sex hindfoot_length
34786     30986     7   1 2000       7         PX                  NA
      weight       genus species   taxa        plot_type
34786     NA Chaetodipus     sp. Rodent Rodent Exclosure
> 
> 
> nrow(surveys)
[1] 34786
> 34786/2
[1] 17393
> survey(17393)
Error in survey(17393) : could not find function "survey"
> survey[17393,]
Error: object 'survey' not found
> surveys[17393,]
      record_id month day year plot_id species_id sex hindfoot_length
17393      9828     1  19 1985      14         AB                  NA
      weight      genus   species taxa plot_type
17393     NA Amphispiza bilineata Bird   Control
> mid_survey <-surveys[17393,]
> mid_survey
      record_id month day year plot_id species_id sex hindfoot_length
17393      9828     1  19 1985      14         AB                  NA
      weight      genus   species taxa plot_type
17393     NA Amphispiza bilineata Bird   Control
> or
surveys[nrow(surveys)/2,]

##creating and standardizing dates
 ymd('1982-1-1')
[1] "1982-01-01"
> paste(surveys$year,surveys$month,surveys$day, sep = "-")
generates a new date colume in the survey
surveys$date<-ymd(paste(surveys$year,surveys$month,surveys$day, sep = "-"))

##Filtering the 129 not parsed
summary(surveys$date)
summary(surveys$date)
        Min.      1st Qu.       Median         Mean      3rd Qu. 
"1977-07-16" "1984-03-12" "1990-07-22" "1990-12-15" "1997-07-29" 
        Max.         NA's 
"2002-12-31"        "129" 
##Adding NAs
> surveys[is.na(surveys$date),c("year","month","day")]
sex<-factor(c("male","female","male","female"))
> levels (sex)
[1] "female" "male"  
> nlevels(sex)
[1] 2
> sex<-factor(sex,levels=c("male", "female"))
> sex
[1] male   female male   female
Levels: male female
> plot(as.factor(surveys$sex))
> levels(surveys$sex)
NULL
> surveys$sex <-addNA(surveys$sex)
> surveys$sex
   [1] M M             M     M M M M   M M M M F F F F F F F M M M F M
  [33] M M F M F F M F F F F F F M F F F F F F M M M F F M F M F F F F
  [65] F M F F F M F M F M M M M M M F M F M F M M M F M F F M F M F F
  [97] F M M M M M M F M F F F F F F F F M M M F F F F F M M M M M M M
 [129] M M M F F F M F M M F M F F F F F F   F M F F M F M M M F F F M
 [161] F F F F F   F F M F F   M F F M F M F F F M F F F F F F F   M M
 [193] F F F F F M M M F F M M M M M M F F M M F F M F M M M M M M M F
 [225] M F M M M F M M M M M F M M F M F M M M F F F F F F F F F F F M
 [257] M F F M F F F F M M F M M F M M F M F F M F F F F F F F F F M F
 [289] F M M F M M M F M M F F F F F F M M M F F M M F M M M M F F M M
 [321] F M M M M M M M M M M M M M M M M F F M M M F F F F M M M F M F
 [353] M M M M M M M F F M M M M F M M M   M F M F M M M F M F M M F M
 [385] F M F M F M F M F M F M M M M M F M M M F F F M F M F M M F M F
 [417] F F F F F M F M F F M F F F M F F F F M F F F F F M M F F F F F
 [449] F F F F F F F F F F F F F F F M F F M M F M F M M F M M F F M F
 [481] M F F M M F M F M M M F M M M M F M F M M M M M M M M F M M F M
 [513] F M F M F F M M M F M M F M M F   F F F M M M M M M F M F M M F
 [545] M M F M F F M F F M M F M M M F F M M M F F F F M F F F M F M M
 [577] M F M M M M M M M M M M M M M M M M M M M M M F M F M M F M F F
 [609] F F F F F M M M M M F M M F M M M F M F F M F M M M F M F F M M
 [641] F F F M M F F M F M M F F M F M M F F F M F M F M M M M F M M M
 [673] M M M M M M M M F M F M M M M F M M M M M M M F M F M M F M M M
 [705] M M M F M M M M F M M M M M F M M M M M M M F M F M M M M M M F
 [737] M F F F F F F F M F M M F M F M M M F M F F M M F F M M F F F M
 [769] M F F F F F F M M M F M   F F F M M M M F F F M M M F F M M   M
 [801] F M F M M M F F M M M M F M M F F F M M M   M F M F M F M M M M
 [833] M M F M F M M F M F M F M M F M F F M M F F M M F M M M M M F M
 [865] M F M F F M F F M M M F M M F M F M M M F M M M F F F F F M F M
 [897] F M F M F M F M F F M M M F F M F F M F F M M F F F M M F M M M
 [929] F M M M M M F M F M F M F M F F F M M M M M F M M M F M M M M F
 [961] F F F M M F F M F F F F M M F M M M F M M M M F M M   M M F F M
 [993] F F F M F F M F
 [ reached getOption("max.print") -- omitted 33786 entries ]
Levels:  F M <NA>
> sex<-as.factor(surveys$sex)
> levels(sex)[1]<- "undetermined"
> plot(sex)

##New command cntrl+shift+m = %>% (pipe)